---
title: "DE3-BRUSH analysis"
author: "cyou"
date: "4/27/2018"
output: html_document
---
DE3: BRUSH Analysis
=================
##### 2018-08-29 Update: After some investigation in PBMC, I think there can be some improvements made in the model for BRUSH
#### some functions for later 
```{r}
## https://gist.github.com/florianhartig/28e06a0ac1fc6d29af3b
source_https <- function(url, ...) {
  # load package
  require(RCurl)
 
  # parse and evaluate each .R script
  sapply(c(url, ...), function(u) {
    eval(parse(text = getURL(u, followlocation = TRUE, cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl"))), envir = .GlobalEnv)
  })
}

## read volcano plot function
source_https("https://raw.githubusercontent.com/kobor-lab/Co-op_Projects/master/ChloeYou/Volcano%20Plot%20with%20CpG%20Labels.R?token=AhczetoVbIlIRM70O93G-Xea5qFWDx9cks5bPTnJwA%3D%3D")
```

## filter out the invariable probes in PBMC 

"To designate a CpG as non-variable in a tissue, a threshold of 5% range in beta values (DNAm level ranging from 0 to 1) between the 10th and 90th percentile was used [16]. While effect sizes as small as 1% are used in EWAS [8, 17, 18], we used a slightly more stringent definition of change in beta of 5% as we are asking only that the population as a whole varies by at least 5% and are not testing an effect size between groups. CpGs with less than 5% reference range of beta values in a single tissue population were considered non-variable in that tissue." [1]

Reference: 

[1] Edgar, R. D., Jones, M. J., Robinson, W. P., & Kobor, M. S. (2017). An empirically driven data reduction method on the human 450K methylation array to remove tissue specific non-variable CpGs. Clinical epigenetics, 9(1), 11.

```{r}
load("adjusted_brush_data.RData")
brush_adj_betas<-adj.residuals_brush

x <- getURL("https://raw.githubusercontent.com/redgar598/Tissue_Invariable_450K_CpGs/master/Invariant_Blood_CpGs.csv")
y <- read.csv(text = x) #114204 
## file y has CpG and RefRange

## finding the DE3 probes that are also in the reference list(overlap) 
DE3_independent_blood_invariable<-betas(de3_bmiq_PBMC)[which(rownames(betas(de3_bmiq_PBMC))%in%y$CpG),]#102297 of the independnt invariable sites are in DE3 PBMC


## Call varibility in DE3
Variation<-function(x) {quantile(x, c(0.9), na.rm=T)[[1]]-quantile(x, c(0.1), na.rm=T)[[1]]}
DE3_ref_range<-sapply(1:nrow(DE3_independent_blood_invariable), function(x) Variation(DE3_independent_blood_invariable[x,]))

## taking out the probes that vary very little(less than 0.05) 
Invariable_in_DE3<-DE3_independent_blood_invariable[which(DE3_ref_range<0.05),]

# Which CpGs are invariable in DE3 and the independent data
invar_in_de3_and_independent<-intersect(y$CpG, rownames(Invariable_in_DE3)) #101821/102297 (99.5%)
DE3_betas_variable<-betas(de3_bmiq_PBMC)[which(!(rownames(betas(de3_bmiq_PBMC))%in%invar_in_de3_and_independent)),]#689498
de3_PBMC_betas<-DE3_betas_variable #689498 that varies 
save(de3_PBMC_betas,file = "de3_PBMC_filtered_invariable_probes.RData")
```

## create M-value file for brush
```{r,eval=FALSE}
setwd("DE3")
load("DE3meta.RData")
load("adjusted_brush_data.RData")
# BRUSH
## take brush meta data from filtermeta 
meta_BRUSH<-filtermeta[filtermeta$SampleType=="BRUSH",]
save(meta_BRUSH,file="BRUSH_meta.RData")
## use adjusted betas for brush
mval_BRUSH<-Mval(brush_adj_betas)
dim(mval_BRUSH) # 791319     50
save(mval_BRUSH,file="mval_BRUSH.RData")

## set filtered air-salin as control group
meta_BRUSH$Exposure<-relevel(meta_BRUSH$Exposure,ref = "FA-S")
```

## split the Mvalue data into three groups, each comparing to the baseline FA-S
```{r}
load("mval_BRUSH.RData")
load("BRUSH_meta.RData")
## extract FA-S, DE-A Mval for linear model
fasdea_sampleID<-meta_BRUSH$SampleID[which(meta_BRUSH$Exposure%in%c("FA-S","DE-A"))]
FASDEA<-mval_BRUSH[,colnames(mval_BRUSH)%in%fasdea_sampleID]
rm(fasdea_sampleID)
## extract FA-S, FA-A Mval for linear model
fasfaa_sampleID<-meta_BRUSH$SampleID[which(meta_BRUSH$Exposure%in%c("FA-S","FA-A"))]
FASFAA<-mval_BRUSH[,colnames(mval_BRUSH)%in%fasfaa_sampleID]
rm(fasfaa_sampleID)
## extract FA-S, PDDE-A Mval for linear model
faspddea_sampleID<-meta_BRUSH$SampleID[which(meta_BRUSH$Exposure%in%c("FA-S","PDDE-A"))]
FASPDDEA<-mval_BRUSH[,colnames(mval_BRUSH)%in%faspddea_sampleID]
rm(faspddea_sampleID)

## checking the data out
dim(FASDEA) # 791319 26  perfectly balanced 
dim(FASFAA) # 791319 24  will remove singled out sample, leaving 11 pairs to analyze
dim(FASPDDEA) # 791319 26 perfectly balanced 
save(FASFAA,FASDEA,FASPDDEA,file = "brush_three_groups_split.RData")

```

side note for running mixed effect models: "ML estimates are unbiased for the fixed effects but biased for the random effects, whereas the REML estimates are biased for the fixed effects and unbiased for the random effects."

## 1. FA-S, DE-A group
```{r,eval=FALSE}
load("CpG_sites_in_order_brush.RData")
load("brush_three_groups_split.RData")
load("BRUSH_meta.RData")


FASDEA_meta<-meta_BRUSH[which(meta_BRUSH$Exposure%in%c("FA-S","DE-A")),]
identical(as.character(FASDEA_meta$SampleID),as.character(colnames(FASDEA))) ## true 

FASDEA_lm<-pbsapply(1:nrow(FASDEA),function(CpG) {
  metaex<-FASDEA_meta
  metaex$Mval<-FASDEA[CpG,]
  mod_brush<-lmer(Mval ~ Exposure+(1|VolunteerNumber),data=metaex)
  mod_brush_2<-lmer(Mval ~ (1|VolunteerNumber),data=metaex)
  anova(mod_brush,mod_brush_2)[2,8]
    }
  )

rm(FASDEA_meta)

## adding rownames and saving file
FASDEA_lm<-as.data.frame(FASDEA_lm)
rownames(FASDEA_lm)<-CpGsites$CpGsites
FASDEA_lm$CpG<-CpGsites$CpGsites
save(FASDEA_lm,file="FASDEA_lm.RData")

## threshold , find hits
hits_FASDEA<-FASDEA_lm[which(FASDEA_lm$FASDEA_lm<=0.0000001),] #  CpG  10^-7
hits_FASDEA_2<-FASDEA_lm[which(FASDEA_lm$FASDEA_lm<=0.000001),] #  CpG 10^-6
hits_FASDEA_3<-FASDEA_lm[which(FASDEA_lm$FASDEA_lm<=0.00001),] #  CpG 10^-5
hits_FASDEA_3<-FASDEA_lm[which(FASDEA_lm$FASDEA_lm<=0.00001),] #  23 CpG 10^-5
hits_FASDEA_4<-FASDEA_lm[which(FASDEA_lm$FASDEA_lm<=0.0001),] #  CpG 10^-4
#hits<- colnames(df)[1:11]
save(hits_FASDEA_3,file="hits_FASDEA.RData")

## creating effect size for FASDEA (DEA-FAS)-> volcano plot
## effect sizes are beta values

FASDEA_beta<-brush_adj_betas[,colnames(brush_adj_betas)%in%FASDEA_meta$SampleID]
dim(FASDEA_beta) #791319     26

## creating a paired dataset to minus the values within pairs omg how do i do this
identical(as.character(colnames(FASDEA_beta)),as.character(FASDEA_meta$SampleID)) # TRUE 

## create two matrix, one for each exposure 
## matrix for DE-A, rownames are SampleID/VolunteerNumber, colnames are CpG sites
FAS_meta<-FASDEA_meta[FASDEA_meta$Exposure=="FA-S",]
DEA_meta<-FASDEA_meta[FASDEA_meta$Exposure=="DE-A",]

DEA_matrix<-FASDEA_beta[,colnames(FASDEA_beta)%in%DEA_meta$SampleID]
FAS_matrix<-FASDEA_beta[,colnames(FASDEA_beta)%in%FAS_meta$SampleID]

colnames(DEA_matrix)<-substr(colnames(DEA_matrix),1,3)
colnames(FAS_matrix)<-substr(colnames(FAS_matrix),1,3)

## order samples to match individual
FAS_matrix<-FAS_matrix[,colnames(DEA_matrix)]

## delta beta is made here
delbeta_matrix<-DEA_matrix-FAS_matrix

FASDEA_delbet<-rowMeans(delbeta_matrix)
FASDEA_delbet<-as.data.frame(FASDEA_delbet)

## make volcano plot and save files
FASDEA_lm<-as.data.frame(FASDEA_lm)
makeVolcano(FASDEA_lm$FASDEA_lm,FASDEA_delbet$FASDEA_delbet,CpGsites$CpGsites,0.05,0.00001,"FASDEA_volcanoplot")
save(FASDEA_lm,delbeta_matrix,FASDEA_delbet,file="FASDEA_effect_size_all_files.RData")

## pull out the hits from volcano plot
FASDEA_delbet$CpG<-rownames(FASDEA_delbet)
 FASDEA_delbet_hits<-FASDEA_delbet[rownames(FASDEA_delbet)%in%rownames(hits_FASDEA_3),]
 FASDEA_delbet_hits<-as.data.frame(FASDEA_delbet_hits)
 FASDEA_delbet_hits<-FASDEA_delbet_hits[order(match(rownames(hits_FASDEA_3),FASDEA_delbet_hits$CpG)),]
 identical(FASDEA_delbet_hits$CpG,rownames(hits_FASDEA_3)) 
FASDEA_hits<-FASDEA_delbet_hits[abs(FASDEA_delbet_hits$FASDEA_delbet)>0.05,] ## this is the object 
rm(FASDEA_delbet_hits)
```

## 2. FA-S, PDDE-A group
```{r,eval=FALSE}
load("CpG_sites_in_order_brush.RData")
load("brush_three_groups_split.RData")
load("BRUSH_meta.RData")


FASPDDEA_meta<-meta_BRUSH[which(meta_BRUSH$Exposure%in%c("FA-S","PDDE-A")),]
identical(as.character(FASPDDEA_meta$SampleID),as.character(colnames(FASPDDEA))) ## true 

FASPDDEA_lm<-sapply(1:nrow(FASPDDEA),function(CpG) {
  metaex<-FASPDDEA_meta
  metaex$Mval<-FASPDDEA[CpG,]
  mod_brush<-lmer(Mval ~ Exposure+(1|VolunteerNumber),data=metaex)
  mod_brush_2<-lmer(Mval ~ (1|VolunteerNumber),data=metaex)
  anova(mod_brush,mod_brush_2)[2,8]
    }
  )

rm(FASPDDEA_meta)

## adding rownames and saving file
FASPDDEA_lm<-as.data.frame(FASPDDEA_lm)
rownames(FASPDDEA_lm)<-CpGsites$CpGsites
FASPDDEA_lm$CpG<-CpGsites$CpGsites
save(FASPDDEA_lm,file="FASPDDEA_lm.RData")

## threshold , find hits
hits_FASPDDEA<-FASPDDEA_lm[which(FASPDDEA_lm$FASPDDEA_lm<=0.0000001),] #   CpG  10^-7
hits_FASPDDEA_2<-FASPDDEA_lm[which(FASPDDEA_lm$FASPDDEA_lm<=0.000001),] #  CpG 10^-6
hits_FASPDDEA_3<-FASPDDEA_lm[which(FASPDDEA_lm$FASPDDEA_lm<=0.00001),] # 17  CpG 10^-5
hits_FASPDDEA_4<-FASPDDEA_lm[which(FASPDDEA_lm$FASPDDEA_lm<=0.0001),] #  CpG 10^-4
save(hits_FASPDDEA_3,file="hits_FASPDDEA.RData")


## creating effect size for FASPDDEA (PDDEA-FAS)-> volcano plot
## effect sizes are beta values
FASPDDEA_beta<-brush_adj_betas[,colnames(brush_adj_betas)%in%FASPDDEA_meta$SampleID]
dim(FASPDDEA_beta) #791319     26

## creating a paired dataset to minus the values within pairs 
identical(as.character(colnames(FASPDDEA_beta)),as.character(FASPDDEA_meta$SampleID)) # TRUE

## create two matrix, one for each exposure 
## matrix for DE-A, rownames are SampleID/VolunteerNumber, colnames are CpG sites
FAS_meta<-FASPDDEA_meta[FASPDDEA_meta$Exposure=="FA-S",]
PDDEA_meta<-FASPDDEA_meta[FASPDDEA_meta$Exposure=="PDDE-A",]

PDDEA_matrix<-FASPDDEA_beta[,colnames(FASPDDEA_beta)%in%PDDEA_meta$SampleID]
FAS_matrix<-FASPDDEA_beta[,colnames(FASPDDEA_beta)%in%FAS_meta$SampleID]

colnames(PDDEA_matrix)<-substr(colnames(PDDEA_matrix),1,3)
colnames(FAS_matrix)<-substr(colnames(FAS_matrix),1,3)

## order samples to match individual
FAS_matrix<-FAS_matrix[,colnames(PDDEA_matrix)]

## delta beta is made here
delbeta_matrix_FASPDDEA<-PDDEA_matrix-FAS_matrix

FASPDDEA_delbet<-rowMeans(delbeta_matrix_FASPDDEA)
FASPDDEA_delbet<-as.data.frame(FASPDDEA_delbet)

## make volcano plot and save files
FASPDDEA_lm<-as.data.frame(FASPDDEA_lm)
makeVolcano(FASPDDEA_lm$FASPDDEA_lm,FASPDDEA_delbet$FASPDDEA_delbet,CpGsites$CpGsites,0.05,0.00001,"FASPDDEA_volcanoplot")
save(FASPDDEA_lm,delbeta_matrix_FASPDDEA,FASPDDEA_delbet,file="FASPDDEA_effect_size_all_files.RData")

## pull out the hits in the volcano plot
FASPDDEA_delbet$CpG<-rownames(FASPDDEA_delbet)
 FASPDDEA_delbet_hits<-FASPDDEA_delbet[rownames(FASPDDEA_delbet)%in%rownames(hits_FASPDDEA_3),]
 FASPDDEA_delbet_hits<-as.data.frame(FASPDDEA_delbet_hits)
 FASPDDEA_delbet_hits<-FASPDDEA_delbet_hits[order(match(rownames(hits_FASPDDEA_3),FASPDDEA_delbet_hits$CpG)),]
 identical(FASPDDEA_delbet_hits$CpG,rownames(hits_FASPDDEA_3)) 
FASPDDEA_hits<-FASPDDEA_delbet_hits[abs(FASPDDEA_delbet_hits$FASPDDEA_delbet)>0.05,] ## this is the object 
rm(FASPDDEA_delbet_hits)
```

## 3. FA-S, FA-A group
this group has only 11 pairs, need to remove 334 and 368 from meta data and beta file
```{r,eval=FALSE}
load("CpG_sites_in_order_brush.RData")
load("brush_three_groups_split.RData")
load("BRUSH_meta.RData")

## remove 334 and 368 from meta data
FASFAA_meta<-meta_BRUSH[which(meta_BRUSH$Exposure%in%c("FA-S","FA-A")),]
FASFAA_meta<-FASFAA_meta[!FASFAA_meta$VolunteerNumber%in%c("334","368"),] ## 22
## remove 334 and 368 from Mval file
FASFAA<-FASFAA[,!colnames(FASFAA)%in%c("334-3","368-4")]
dim(FASFAA) #791319     22

identical(as.character(FASFAA_meta$SampleID),as.character(colnames(FASFAA))) ## true 

FASFAA_lm<-pbsapply(1:nrow(FASFAA),function(CpG) {
  metaex<-FASFAA_meta
  metaex$Mval<-FASFAA[CpG,]
  mod_brush<-lmer(Mval ~ Exposure+(1|VolunteerNumber),data=metaex)
  mod_brush_2<-lmer(Mval ~ (1|VolunteerNumber),data=metaex)
  anova(mod_brush,mod_brush_2)[2,8]
    }
  )

rm(FASFAA_meta)

## adding the list of CpG names into the pval file
FASFAA_lm<-as.data.frame(FASFAA_lm)
rownames(FASFAA_lm)<-CpGsites$CpGsites
FASFAA_lm$CpG<-CpGsites$CpGsites
save(FASFAA_lm,file="FASFAA_lm.RData")

## threshold , find hits
hits_FASFAA<-FASFAA_lm[which(FASFAA_lm$FASFAA_lm<=0.0000001),] #   CpG  10^-7
hits_FASFAA_2<-FASFAA_lm[which(FASFAA_lm$FASFAA_lm<=0.000001),] # 3 CpG 10^-6
hits_FASFAA_3<-FASFAA_lm[which(FASFAA_lm$FASFAA_lm<=0.00001),] #  31 CpG 10^-5
hits_FASFAA_4<-FASFAA_lm[which(FASFAA_lm$FASFAA_lm<=0.0001),] # 

save(hits_FASFAA_3,file="hits_FASFAA.RData")

## creating effect size for FASFAA (FAA-FAS)-> volcano plot
## effect sizes are beta values
FASFAA_beta<-brush_adj_betas[,colnames(brush_adj_betas)%in%FASFAA_meta$SampleID]
dim(FASFAA_beta) #791319     22

## creating a paired dataset to minus the values within pairs
identical(as.character(colnames(FASFAA_beta)),as.character(FASFAA_meta$SampleID)) # TRUE

## create two matrix, one for each exposure 
## matrix for DE-A, rownames are SampleID/VolunteerNumber, colnames are CpG sites
FAS_meta<-FASFAA_meta[FASFAA_meta$Exposure=="FA-S",]
FAA_meta<-FASFAA_meta[FASFAA_meta$Exposure=="FA-A",]

FAA_matrix<-FASFAA_beta[,colnames(FASFAA_beta)%in%FAA_meta$SampleID]
FAS_matrix<-FASFAA_beta[,colnames(FASFAA_beta)%in%FAS_meta$SampleID]

colnames(FAA_matrix)<-substr(colnames(FAA_matrix),1,3)
colnames(FAS_matrix)<-substr(colnames(FAS_matrix),1,3)

## order samples to match individual 
FAS_matrix<-FAS_matrix[,colnames(FAA_matrix)]

## delta beta is made here
delbeta_matrix_FASFAA<-FAA_matrix-FAS_matrix

FASFAA_delbet<-rowMeans(delbeta_matrix_FASFAA)
FASFAA_delbet<-as.data.frame(FASFAA_delbet)

## make volcano plot and save files
FASFAA_lm<-as.data.frame(FASFAA_lm)
makeVolcano(FASFAA_lm$FASFAA_lm,FASFAA_delbet$FASFAA_delbet,CpGsites$CpGsites,0.05,0.00001,"FASFAA_volcanoplot")
save(FASFAA_lm,delbeta_matrix_FASFAA,FASFAA_delbet,file="FASFAA_effect_size_all_files.RData")

## pull out the hits in the volcano plot
FASFAA_delbet$CpG<-rownames(FASFAA_delbet)
 FASFAA_delbet_hits<-FASFAA_delbet[rownames(FASFAA_delbet)%in%rownames(hits_FASFAA_3),]
 FASFAA_delbet_hits<-as.data.frame(FASFAA_delbet_hits)
 FASFAA_delbet_hits<-FASFAA_delbet_hits[order(match(rownames(hits_FASFAA_3),FASFAA_delbet_hits$CpG)),]
 identical(FASFAA_delbet_hits$CpG,rownames(hits_FASFAA_3)) 
FASFAA_hits<-FASFAA_delbet_hits[abs(FASFAA_delbet_hits$FASFAA_delbet)>0.05,] ## this is the object 
rm(FASFAA_delbet_hits)

```

## merge all hits into one data set 
```{r}
# merge the three hit data frames, leaving three columns:CpG,Group,Pval
## rename Pval column
colnames(FASDEA_hits)[colnames(FASDEA_hits)=="FASDEA_delbet"] <- "effect size"
colnames(FASFAA_hits)[colnames(FASFAA_hits)=="FASFAA_delbet"] <- "effect size"
colnames(FASPDDEA_hits)[colnames(FASPDDEA_hits)=="FASPDDEA_delbet"] <- "effect size"
## add Group Column 
FASDEA_hits$Group<-"FASDEA"
FASFAA_hits$Group<-"FASFAA"
FASPDDEA_hits$Group<-"FASPDDEA"

## combine data sets
hits_brush_vol<-rbind(FASFAA_hits,FASDEA_hits,FASPDDEA_hits)
save(FASDEA_hits,FASFAA_hits,FASPDDEA_hits,hits_brush_vol,file="volcano_hits_brush.RData")


# merge the three hit data frames, leaving three columns:CpG,Group,Pval
## rename Pval column
colnames(hits_FASDEA_3)[colnames(hits_FASDEA_3)=="FASDEA_lm"] <- "Pval"
colnames(hits_FASFAA_3)[colnames(hits_FASFAA_3)=="FASFAA_lm"] <- "Pval"
colnames(hits_FASPDDEA_3)[colnames(hits_FASPDDEA_3)=="FASPDDEA_lm"] <- "Pval"
## add Group Column 
hits_FASDEA_3$Group<-"FASDEA"
hits_FASFAA_3$Group<-"FASFAA"
hits_FASPDDEA_3$Group<-"FASPDDEA"

## combine data sets
hits_brush<-rbind(hits_FASDEA_3,hits_FASFAA_3,hits_FASPDDEA_3)

## save files 
save(hits_FASDEA_3,hits_FASFAA_3,hits_FASPDDEA_3,hits_brush,file="all_hits_brush.RData")
```

## create plots to analyze the hits found in each group
### 1. boxplot
for each hit, create a boxplot with x-axis being the two seperate exposures and y-axis being the beta values
```{r}
# use "hits_brush" file

## comment out the code below, found it online 
## ggplot(mtcars,aes(factor(cyl),mpg))+geom_boxplot()+
##   geom_point(aes(color=factor(am)),position=position_dodge(width=0.5))

## I guess i need to create a new file with the beta value of one CpG site from all 26/24 samples, 13 in each side of the boxplot

## create new file for CpG site boxplot
## extract rows(CpG) from the brush_adj_beta
compiled_hits<-brush_adj_betas[rownames(brush_adj_betas)%in%hits_brush$CpG,]
dim(compiled_hits) # 71 50

## makes rows into SampleID and columns into CpG
compiled_hits<-as.data.frame(t(compiled_hits))
## check out compiled_hits

## add another column of Exposure group
identical(as.character(rownames(compiled_hits)),as.character(colnames(brush_adj_betas))) #true

## create ggplots 
## for each hit, we will plot all samples, including the exposures that weren't associated with finding this hit 

compiled_hits<-brush_adj_betas[rownames(brush_adj_betas)%in%hits_brush_vol$CpG,]
dim(compiled_hits) # 71 50
compiled_hits<-as.data.frame(t(compiled_hits))
identical(as.character(rownames(compiled_hits)),as.character(colnames(brush_adj_betas))) #true
## add another column to "compiled_hits" indicating the exposure of the sample 
identical(as.character(rownames(compiled_hits)),as.character(meta_BRUSH$SampleID)) #true
compiled_hits$Exposure<-meta_BRUSH$Exposure
compiled_hits$VolunteerNumber<-substr(rownames(compiled_hits),1,3)

## general boxplot
my_comparisons <- list( c("FA-S", "DE-A"), c("FA-S", "FA-A"), c("FA-S", "PDDE-A"))
plots <- lapply(1:7, function(i) ggplot(compiled_hits,aes(as.factor(compiled_hits$Exposure),compiled_hits[,i],color=as.factor(compiled_hits$Exposure)))+geom_boxplot()+geom_point(shape=19,size=1, position=position_jitter(w=0.25),color="grey")+theme_bw()+xlab("Exposure")+ylab("Beta Value")+labs(title=colnames(compiled_hits)[i])+ stat_compare_means(comparisons = my_comparisons,method = "wilcox.test"))

do.call(grid.arrange, plots) 
```

### 2. spaghetti plot for the three paired groups 
```{r}
## use "compiled_hits"
## creating 7*3-2(missing data) plots

seven<-list()
for(i in 1:7){
 newdf<-compiled_hits[,c(i,8,9)]
 FAA<-newdf[newdf$Exposure%in%c("FA-S","FA-A"),] # 24
 FAA<-FAA[!FAA$VolunteerNumber%in%c("368","341"),]
 DEA<-newdf[newdf$Exposure%in%c("FA-S","DE-A"),] # 26
 PDDEA<-newdf[newdf$Exposure%in%c("FA-S","PDDE-A"),] # 26
 
 FAA$Exposure<- factor(FAA$Exposure, levels=c("FA-S","FA-A"))
 DEA$Exposure<- factor(DEA$Exposure, levels=c("FA-S","DE-A"))
 PDDEA$Exposure<- factor(PDDEA$Exposure, levels=c("FA-S","PDDE-A"))
 
  p1<-ggplot(FAA,aes(as.factor(FAA$Exposure),FAA[,1],color=as.factor(FAA$VolunteerNumber),group=FAA$VolunteerNumber))+geom_line()+geom_point(shape=19,size=1,color="grey")+theme_minimal()+xlab("Exposure")+ylab("Beta Value")+labs(title=colnames(FAA)[1])+theme(legend.position='none')
 
  p2<-ggplot(DEA,aes(as.factor(DEA$Exposure),DEA[,1],color=as.factor(DEA$VolunteerNumber),group=DEA$VolunteerNumber))+geom_line()+geom_point(shape=19,size=1,color="grey")+theme_minimal()+theme(legend.position='none')+labs(x="", y="")
  
  p3<-ggplot(PDDEA,aes(as.factor(PDDEA$Exposure),PDDEA[,1],color=as.factor(PDDEA$VolunteerNumber),group=PDDEA$VolunteerNumber))+geom_line()+geom_point(shape=19,size=1,color="grey")+theme_minimal()+theme(legend.position='none')+labs(x="", y="")
   
  seven[[i]]<-grid.arrange(p1,p2,p3,ncol=3)
}

do.call("grid.arrange", c(seven, ncol=2))

```

#### check UCSC genes on the 7 hits found in the volcano plots
```{r}
fd<-fData(de3_bmiq_BRUSH)
fd<-fd[rownames(fd)%in%hits_brush_vol$CpG,]  # 7
colnames(fd) #"UCSC_REFGENE_NAME" idk what other columns i need to pay attention to
UCSC_gene<-as.data.frame(fd$UCSC_REFGENE_NAME)
```
#### limitations: deconvolution method; the model can be better; lots of invariable probes;...
